The Courtroom Challenge—A Conjecture on the Unavoidable Cost of LLM Verification

### 2025-06-30 paper/main.tex
Add files via upload

To address these shortcomings we pose the \textit{Courtroom Challenge}: a
hybrid state machine in which every LLM claim is tested against a persistent
precedent store and—when automatic checks falter—escalated under a
tunable \textit{validation\_preference} dial (formal spec in
Section~\ref{sec:framework}).  The design rewires known parts
(generator, retrieval, human oversight) but prices each byte, cycle,
and minute of validation.  Retrieval-augmented generation (RAG) provides the raw materials
\citep{lewis2020rag,fan2024ragsurvey}.  
Recent advances—Self-RAG's self-verification \citep{asai2023selfrag}, 
Iter-RetGen's retrieval-generation synergy \citep{shao2023iterretgen}, 
MemoRAG's compressed long-term memory \citep{qian2024memorag}, and 
Chain-of-Verification RAG's answer-driven re-checks \citep{he2024covrag}—motivate 
our persistent precedent store but still validate only per answer.  
Human-alignment work such as Constitutional AI 
\citep{bai2022constitutional} motivates the escalation branch.  
Whereas MemoRAG retains past evidence without auditing it, our buffer makes
each accepted verdict both permanent and costly.
Unlike debate frameworks \citep{chen2024reconcile,madaan2023selfrefine} or formal
provers \citep{wang2025malot}, we emphasise continuous cross-session
verification: knowledge is never final, only current.

@@

  \item \textbf{Memory update rule.}
  \(U : (M, q, h, a) \to M'\) records (or merges) a precedent whenever a claim is
  accepted.  Whereas vanilla RAG treats retrieved passages as a transient
  context window \citep{lewis2020rag}—and even iterative or self-reflective
  extensions still validate only the current answer
  \citep{shao2023iterretgen,asai2023selfrag}—the Courtroom store is
  persistent: it survives across sessions and every entry can later be
  re-audited for contradiction or obsolescence.

@@

Large-language-model assistance: Some portions of the manuscript—
in particular early wording suggestions, reference formatting, and a toy-demo
code sketch—were drafted or refined with the aid of OpenAI ChatGPT
(model o3, May 2025). All ideas, experiments, and final text were reviewed,
verified, and are the sole responsibility of the author.

>>>>

To address these shortcomings we pose the \textit{Courtroom Challenge}: a
hybrid state machine in which every LLM claim is tested against a persistent
precedent store and—when automatic checks falter—escalated under a
tunable \textit{validation\_preference} dial (formal spec in
Section~\ref{sec:framework}).  The design rewires known parts
(generator, retrieval, human oversight) but prices each byte, cycle,
and minute of validation.  Retrieval-augmented generation (RAG) provides the raw materials
\citep{lewis2020rag,fan2024ragsurvey}.  
Recent advances—Self-RAG's self-verification \citep{asai2024selfrag}, 
Iter-RetGen's retrieval-generation synergy \citep{shao2023iterretgen}, 
MemoRAG's compressed long-term memory \citep{qian2024memorag}, and 
Chain-of-Verification RAG's answer-driven re-checks \citep{he2024covrag}—motivate 
our persistent precedent store but still validate only per answer.  
Human-alignment work such as Constitutional AI 
\citep{bai2022constitutional} motivates the escalation branch.  
Whereas MemoRAG retains past evidence without auditing it, our buffer makes
each accepted verdict both permanent and costly.
Unlike debate frameworks \citep{chen2024reconcile,madaan2023selfrefine} or formal
provers \citep{wang2025malot}, we emphasise continuous cross-session
verification: knowledge is never final, only current.

@@

  \item \textbf{Memory update rule.}
  \(U : (M, q, h, a) \to M'\) records (or merges) a precedent whenever a claim is
  accepted.  Whereas vanilla RAG treats retrieved passages as a transient
  context window \citep{lewis2020rag}—and even iterative or self-reflective
  extensions still validate only the current answer
  \citep{shao2023iterretgen,asai2024selfrag}—the Courtroom store is
  persistent: it survives across sessions and every entry can later be
  re-audited for contradiction or obsolescence.

@@

Large-language-model assistance: Some portions of the manuscript—
in particular early wording suggestions, reference formatting, and a toy-demo
code sketch—were drafted or refined with the aid of OpenAI ChatGPT
(model o3, June 2025). All ideas, experiments, and final text were reviewed,
verified, and are the sole responsibility of the author.
